Secure Composition of Robust and Optimising Compilers

### 2023-07-11 popl/main.tex
upto ex 3.8

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets}\MP{should cite catalin's MS paper too}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, property satisfaction is defined as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}\MP{many more things on RS here}, the concept of satisfaction can be extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity}, \MP{Add papers from Amal. Papers from just Dominique on FAC. RS works.}
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey}\MP{cite kennedy and teh general SC dagstuhl and ACM papers too}, and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.

For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.
\MPin{
	somewhere we should say that a class of props can also include a single prop
}

@@

\begin{example}[Types]
  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
  Consider the following $\src{L}$ component and its compiled version below.
  \begin{center}
  $
  \begin{array}{l}
    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
  \end{array}
  $
  \end{center}
  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
  Thus, even if $\src{foo}$ may have been robust with respect to the no-reads-out-of-bounds property\MP{sms?}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
\end{example}

@@

This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
Then, this section introduces monitors for each of the properties presented earlier (\Cref{subsec:monitors}).
\MP{why?}

@@

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
The purpose of the model is to define key security properties of interest, such as \gls{ms} or a stricter variant of cryptographic constant time.
To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}~\cite{nagarakatte2010cets}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone are enough to provide a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

\gls{sms} disregards\MP{prohibits?} out-of-bounds accesses:

@@

Full \gls{ms} is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}
\MPin{
	cite catalin's paper and the mswasm paper for the def of ms
}

@@

\gls{cct} is a hyperproperty\MP{hypersafety?}\MP{what is a hp?} and, thus, not as easy to check with just monitors.\MP{chy? cit?} %Beutner et al
To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace.\MP{cit? did we take some inspiration}

@@

The combination of \gls{ms} and \gls{scct} is the intersection of these properties, \gls{msscct}.
It entails all program behaviours that are both memory safe and strictly cryptographic constant time.

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\left\{\trace \left| \begin{array}{rcl}
                                      \trace\in\mssafe & \text{ and } & \trace\in\scctsafe
                                    \end{array}\right.\right\}
  $
\end{definition}
\MPin{
	comment on the composition at this levle?
}

@@

Monitors can be used to enforce safety properties and this section presents concrete instantiations of monitors for the previousvy presented properties (\Cref{subsec:propdefs}).
\MPin{
	why
}
To lessen the burden when proving\MP{proving what}, each monitor makes use\MP{uses} of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor-level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation projects the events\MP{whaddaya mean?}, while the stutter-equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasise the kind of event agreement.\MP{unclear sentence}
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

For \gls{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
In the reflexive case of $\monitorcheck{\tmsmonitor}{\tmsmonitor}{\absevent}$, the monitor $\tmsmonitor$ must not contain allocated locations, i.e., $\tmsmonitor.allocated=\emptyset$.\MP{wut?}
This ensures that programs do not forget to free memory.

@@

\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmsafe}
  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
\end{lemma}

\begin{example}[A program satisfying \gls{tms}]
  Consider the following program which, when run, yields the trace $\ev{Alloc}\ \loc[_{1}]\ 17;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Use}\ \loc[_{1}]\ 5;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{1}];\comp;\unlock$, where $\loc[_{1}]$ is some memory address determined by the semantics of the programming language.
  \begin{lstlisting}[language=c,basicstyle=\small\ttfamily,morekeywords={new,delete}]
int main() {  // allocates 17 integers initialised to 2
  int* x = new[17](2);

  x[5] = 12;
  int y = x[5];
  delete x;

  return y;
}
  \end{lstlisting}
  Relating the trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Use}\ \loc[_{1}]\ \comp\cdot\absev{Dealloc}\ \loc[_{1}]\ \comp$.
  It is easy to see that $\montmssafe{\abstrace}$.
  From \Thmref{lem:mon:tmsafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}.
\end{example}
\MPin{
	example is good, but it'd be nice to tie it to the running example?
}

>>>>

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid~\cite{clarkson2008hyper}.
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}.
A class is simply a set of properties and for the class of safety properties, it is decidable whether a trace satisfies a safety property with just a finite trace prefix.
As an example, consider a trace describing an interaction with a memory where the deallocation of an address $\loc$ precedes a read at that address in memory: $\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$.
This program behaviour is insecure with respect to a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls{tms} and there is no way to append events to this prefix which would result in the trace being admissible.
In the following, the execution of a whole program $\wholeprogvar$ that terminates in state $\runtimetermvar$ according to the language semantics and produces trace $\trace$ is written as $\progstepto{\wholeprogvar}{\runtimetermvar}{\trace}$.
With this, property satisfaction is defined as follows:
\bul{whole programs $\wholeprogvar$ satisfy a property $\pi$} iff \iul{$\wholeprogvar$ yields a trace $\trace$} such that \oul{$\trace$ satisifies $\pi$} (\Cref{def:propsat}).

@@

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,patrignani2021rsc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps}, the concept of satisfaction can be extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.

For \bul{compilers $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of properties $\class$}, if for any \rul{property $\pi$ of that class $\class$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\pi$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\pi$}.

@@

Note that a class of properties $\class$ can represent just one property $\pi$ by lifting~\cite{clarkson2008hyper} that property to sets of properties, i.e., use the powerset of $\pi$ instead of $\pi$ itself.
Because of this, this paper writes $\rtp{\cc{\src{L}}{\trg{L}}}{\pi}$, even though $\pi$ is a property and not a class.

\begin{example}[Types]
  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
  Consider the following $\src{L}$ component and its compiled version below.
  \begin{center}
  $
  \begin{array}{l}
    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
  \end{array}
  $
  \end{center}
  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
  Thus, even if $\src{foo}$ may have been robust with respect to the \gls{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Thmref{def:rtp}.
\end{example}

@@

This section introduces a trace model and uses it to define the key properties of interest for this paper: \gls{tms}, \gls{sms}, \gls{ms}, and \gls{scct} (\Cref{subsec:propdefs}).
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study (\Cref{sec:casestud:defs,sec:casestud:rtp}) this paper presents later. 
Lastly, for each of the key properties, this section introduces corresponding monitors (\Cref{subsec:monitors}) that check them.

@@

The specification trace model defines events as either the empty event ($\emptyevent$), a crash ($\lightning$), or as tuples consisting of a pre-event, a control-tag, and a security-tag.
The purpose of the model is to define key security properties of interest, such as \gls{ms} or a stricter variant of cryptographic constant time.
To this end, security-tags indicate whether an event contains sensitive information ($\lock$) or not ($\unlock$), while control-tags state whether the context ($\ctx$) or the component ($\comp$) are responsible for emitting the event.
The latter is necessary to be able to ignore actions done by a spurious context that, e.g., immediately deallocates a memory location twice, thus violating \gls{tms}~\cite{nagarakatte2010cets}.
%Even though this could be used for blame~\cite{patrignani2023blame}, this is not what this paper is concerned about.
Lastly, pre-events describe the actual kind of event that happened.
One such kind is the allocation event ($\ev{Alloc\ \loc\ n}$) that fires whenever a program claims $n$ cells of memory and stores them at address $\loc$.
Dually, deallocation ($\ev{Dealloc\ \loc}$) announces that the object at location $\loc$ is freed.
These two events alone are enough to provide a partial description of \gls{tms} by requiring that, e.g., there is only one deallocation event that carries a location $\loc$.
To be able to express \gls{sms}, there is also an event to describe reads from and writes to memory ($\ev{Use\ \loc\ n}$).
Finally, for cryptographic code, there is a general guideline that secrets must not be visible on a trace.
Moreover, an instruction whose timing is data-dependent must not have a secret as an operand.
Typical operations with data-dependent timing are branches and certain binary operations, such as division\footnote{This is highly architecture-dependent, but division is an operation that serves as a classic example for a data-dependent timing instruction, e.g., \cite[p.~755]{arm-refman}.}.
Both operations are also modelled in the specification trace model ($\ev{Branch\ n}$ and $\ev{Binop\ n}$).

@@

\gls{sms} prohibits out-of-bounds accesses:

@@

Full \gls{ms} (similar to earlier work~\cite{nagarakatte2009soft,nagarakatte2010cets,jim2002cyclone,necula2005ccured,michael2023mswasm}) is then described as the conjunction of \Cref{def:trace:tmsdef,def:trace:smsdef}.
%%% Catalin's paper doesn't apply, since their definition of MS relies on noninterference to guarantee data isolation.
%%% our definition does not guarantee that (on purpose, we want to compose it with cct)
Note, however, that this definition says nothing about memory-safety issues introduced by side-channels, such as speculation.

\begin{definition}[\glsfirst{ms}]\label{def:trace:msdef}
  $
  \mssafe:=\left\{\trace \left|\begin{array}{rcl}
                                 \trace\in\tmssafe & \text{ and } & \trace\in\smssafe
                               \end{array}\right.\right\}
  $
\end{definition}

@@

\gls{cct} is a hypersafety property~\cite{barthe2018sec} and, thus, difficult to check with monitors.
This is because, intuitively, hypersafety properties can relate multiple execution traces with eachother, but monitors work on a single execution. not as easy to check with just monitors.
To remedy this, this section defines the property \gls{scct}, a stricter variant of \gls{cct} that enforces the policy that no secret should appear on a trace (inspired by earlier work~\cite{almeida2017jasmin}).

@@

The combination of \gls{ms} and \gls{scct} is the intersection of these properties, \gls{msscct}.
Since \gls{ms} and \gls{scct} are just sets of traces that, intuitively, contain all program behaviors that follow a security policy, the intersection of them contains all program behaviors that follow both security policies, i.e., it entails all program behaviours that are both \gls{ms} and \gls{scct}.

\begin{definition}[\glsfirst{ms} and \gls{scct}]\label{def:trace:msscctdef}
  $
  \msscctsafe:=\mssafe\cap\scctsafe
  $
\end{definition}

@@

\begin{figure}[h!]
  \judgbox{\traceagree{\pi}{\trace}{\abstrace}}{,,Specification-level trace $\trace$ agrees with monitor-level trace $\abstrace$ with respect to property $\pi$.''}
  %
  \typerule{traceagree-empty}{}{
    \traceagree{\pi}{\hole{\cdot}}{\hole{\cdot}}
  }{traceagree-empty}
  %
  \typerule{traceagree-ign-L}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\emptyevent\cdot\trace}{\abstrace}
  }{traceagree-ign-L}
  %
  \typerule{traceagree-ign-R}{
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\trace}{\absev{\emptyevent}\cdot\abstrace}
  }{traceagree-ign-R}
  %
  \typerule{traceagree-cons}{
    \traceagree[]{\pi}{\event}{\absevent} &
    \traceagree{\pi}{\trace}{\abstrace}
  }{
    \traceagree{\pi}{\event\cdot\trace}{\absevent\cdot\abstrace}
  }{traceagree-cons}
  \caption{Trace-Agreement relation that equates specification-level traces with monitor-level traces.}\label{fig:stuttereq}
\end{figure}

Monitors enforce safety properties by accepting or rejecting traces, i.e., if it rejects a trace, the trace does not satisfy the property the monitor checks.
Since reasoning on monitors is easier than directly on just traces, this section presents a monitor for each of the previously shown safety properties (\Cref{subsec:propdefs}).
To lessen the burden when proving that a monitor accepts the trace of a program execution, each monitor uses a custom trace model that contains only the relevant information related to the property the monitor checks.
To go from specification traces $\trace$ to monitor-level traces $\abstrace$, each property $\pi$ has an associated event agreement relation $\traceagree[]{\pi}{\event}{\absevent}$.
\Cref{fig:stuttereq} shows how the event agreement is lifted to traces.
The trace agreement is the same for all properties $\pi$ up to the use of the event agreement in \Cref{tr:traceagree-cons}.
With agreements, this section defines monitor satisfaction for traces and then it proves that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor-satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

For \gls{tms}, the state of the monitor is a record with two sets keeping track of allocated and deallocated locations.
%The monitor runs on a modified trace semantics with more abstract $\absev{Alloc\ }\loc$,$\absev{Dealloc\ }\loc$,$\absev{Use\ }\loc$, and $\absterm$ events, which represent {\em allocation}, {\em deallocation}, and {\em use} of a location $\loc$ as well as abnormal program termination.
\Cref{tr:tms-use} simply requires that a location is (i) allocated and (ii) not freed.
\Cref{tr:tms-alloc,tr:tms-dealloc} both require a location to not be freed already and extend the monitor state accordingly.
This restriction effectively disallows reallocation to reassign the same location to an object.
However, the definition can easily be adapted by, e.g., attaching a natural number serving as a counter.
Contrary to other monitors in this paper, the multi-step relation of the \gls{tms} monitor is non-standard:
\begin{center}
  \judgbox{\monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}}{,,Monitor $\tmsmonitor$ multi-steps to $\tmsmonitor[']$ given trace $\abstrace$.''}$\;$\\
  %
  \typerule{tms-refl}{
    \tmsmonitor.\text{allocated}=\emptyset
  }{
    \monitorcheck{\tmsmonitor}{\tmsmonitor}{\hole{\cdot}}
  }{tms-refl}
  %
  \typerule{tms-ign-trans}{
    \monitorcheck{\tmsmonitor}{\tmsmonitor[']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{\emptyevent}\cdot\abstrace}
  }{tms-ign-trans}
  %
  \typerule{tms-trans}{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}
    \monitorcheck{\tmsmonitor[']}{\tmsmonitor['']}{\abstrace}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor['']}{\absevent\cdot\abstrace}
  }{tms-trans}
\end{center}
\Cref{tr:tms-ign-trans,tr:tms-trans} are the same for all monitors, but \Cref{tr:tms-refl} has, in this case, an additional premise that no more locations should be allocated.
This rejects the behavior of programs that forget to free memory.

@@

\begin{lemma}[Traces with Monitor Satisfaction are $\tmssafe$]\label{lem:mon:tmssafe}
  If $\montmssafe{\trace}$, then $\trace\in\tmssafe$. %\Coqed
\end{lemma}

\begin{example}[A program not satisfying \gls{tms}]
  Consider the following \texttt{C++11} program that calls \texttt{strncpy} (\Cref{ex:strncpy}) and prints the result to the standard output stream.

  \begin{lstlisting}[language=c++,basicstyle=\small\ttfamily,mathescape,commentstyle=\color{gray}]
int main() {  // allocates 12 chars containing a greeting message
  char* greetings = new char[12] { "Hello POPL!" }; // <- address ${\color{gray}\loc_{x}}$
  char* to = new char[12];                          // <- address ${\color{gray}\loc_{y}}$

  strncpy(12, to, greetings);
  delete to;

  printf("%cOPL\n", to[6]);
}
  \end{lstlisting}
  Up to the body of \texttt{printf}, the program execution yields the specification trace $\ev{Alloc}\ \loc[_{x}]\ 12;\comp;\unlock\cdot\ev{Alloc}\ \loc[_{y}]\ 12;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 0;\comp;\unlock\cdot\ev{Use}\ \loc_{x}\ 1;\comp;\unlock\cdot\ldots\cdot\ev{Use}\ \loc_{x}\ 12;\comp;\unlock\cdot\ev{Dealloc}\ \loc[_{y}];\comp;\unlock\cdot\ev{Use}\ \loc_{y}\ 6;\comp;\unlock$.
  Relating this trace to abstract monitor events yields $\abstrace=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  Remembering the definition of \texttt{strncpy} (\Cref{ex:strncpy}), it is easy to see that it does not deallocate its arguments.
  Even though the trace contains an out-of-bounds access right before returning from \texttt{strncpy}, this is no concern for \gls{tms}, since the location $\loc[_{x}]$ is still allocated.
  However, having returned from \texttt{strncpy}, the \texttt{main} function continues and deallocates $\loc[_{y}]$ whose subsequent use in the \texttt{printf} call is a use-after-free bug.
  The fix would be to \lstinline[language=c++,basicstyle=\small\ttfamily]|delete greetings| instead of \texttt{to} and add a \lstinline[language=c++,basicstyle=\small\ttfamily]|delete to| after the \texttt{printf} call, which leads to the abstract monitor trace $\abstrace[']=\absev{Alloc}\ \loc[_{x}]\ \comp\cdot\absev{Alloc}\ \loc[_{y}]\ \comp\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp\cdot\ldots\cdot\absev{Use}\ \loc[_{x}]\ \comp\cdot\absev{Dealloc}\ \loc[_{x}]\ \comp\cdot\absev{Use}\ \loc[_{y}]\ \comp$.
  It is easy to see that $\montmssafe{\abstrace[']}$ and from \Thmref{lem:mon:tmssafe}, it follows that the program satisfies \Thmref{def:trace:tmsdef}, even though the program still violates \gls{sms}.
\end{example}
